---
title: "Lab 6"
author: "Jack Miller"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rand based test example

Mann-Whitney U (non-parametric alternatibe to an unpaired 2-sample t-test). 

H0: Equal ranks (or medians) - this is the null hypothesis
HA: Ranks are not equal (or, medians are not equal) - this is the laternative hypothesis

## Create samples

'set.seed() to create "pseudoradom, but reproducible samples"
      if we all have the same starting point, then we should all have the same sample
      
```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace  = TRUE) # 20 = the largest vale, 15 = # of outputs; replace = if values can repeat

gp_1

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE) 
gp_2
```

Take a quick look at the distributions sing 'hist()'

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If thee data were paired, I can just add the argument 'paired = 'TRUE"' within the t.test() function.
# t-test is a question about mean differences 
# the p-value means that thatis the prob of finding 2 samples with 2 means that are least this difference by random chance, if  they were drawn from populations with the same mean
# This 
ttest
```

This p-value means that if our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is r round(ttest$p.value.2)' probability of finding sample means *at least as different* as mine by random chance. 

[A more comprehensive explanation of difference between groups... actually means, data spread, effect size]... there is no significant difference (t(df)= t-statistic value, p = ). Degrees of freedom is a measure of ???

### Mann Whitney U rank based test 

We use the 'wilcox.test()' function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
mwu
```

This p-value means that if our samples are drawn from a populations with the same median, there is a 28% chance that we will find sample medians that are at least as different as ours by random chance. 

NO significant different in medians (Mann-Whitney U; U = 'r mwu$statistic', p = 'r round(mwu$p.value, 2)'). 

If you want to check on the components of a function, use "function_name" in Console and check on Help to the right of the page.

Example text for reporting: "The medians of group A (M = 22.4g) and group B(m=21.2 g) did not differ significantly (Mann Whitney U, ...)."

## Simple linear regression 

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense. 

Use the 'penguins' data from (palmerpenguins), exploring flipper length (x) versus body mass (y). 

```{r}
# Use view(penguins) to look at it 
# Make an exploratory scatter plot of penguin flipper length (x-axis) versus body mass (y-axis).

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


### Find the model estimates (coefficients)

Use the 'lm()' function for linear model.

body_mass_g ~ <<<< tilde 

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
penguin_lm
# body mass (g) = 49.69 (flipper length (mm)) - 5780 (the intercept doesn't matter because you cannot have a penguin of weight = 0 g)
# on average, for 1 mm increase in flipper length, it is associated with a 49.69 increase in g 
# you can do summary(penguin_lm) to find 
```


Model coefficients 'r penguin_lm$coefficients'
flipper length coefficient: 'r prenguin_lm$coefficiencts[2]'

### Use 'broom::tidy()' to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```
### Use broom:glance() to get output of model fit & statistical significance 

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance 

penguin_lm_tidy$estimate[2]
```

### Use equatiomatic::extract_eq() to return model in LaTeX 

LaTeX is a language use to write math expressions in plain text

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use plot() function to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```





